---
title: "R Notebook"
output: html_notebook
---
Data Cleaning & Visualization(Outliers)

```{r}
setwd("C:/Users/Edwyn/R")
column.labels <- c("Region","Alcohol","Malic Acid","Ash","Alk","Magn","TotalPhenoles","Flavonoids","NonFL","Proanthocyanins","ColorInt","Hue","OD","Proline")

wine <- read.csv("wine.csv", header = F)
colnames(wine) <- column.labels

wine.N <- subset(wine, select = -c(Region))

#Inspect
summary(wine.N)
```

Method 1 (Outlier Detector) - Bivariate Box Plot
```{r}
library(MVA)
bvbox(wine.N[,c(1,7)], pch ="", xlab = "Alcohol", ylab = "Flavonoids", main = "Bivariate Boxplot of Alcohol vs Flavonoids")
text(wine.N$Alcohol, wine.N$Flavonoids, labels = rownames(wine.N), cex = 0.7)

bvbox(wine.N[,c(3,4)], pch ="", xlab = "Ash", ylab = "Alk", main = "Bivariate Boxplot of Ash vs Alk")
text(wine.N$Ash, wine.N$Alk, labels = rownames(wine.N), cex = 0.7)

bvbox(wine.N[,c(11,10)], pch ="", xlab = "Hue", ylab = "ColorInt", main = "Bivariate Boxplot of Hue vs ColorInt")
text(wine.N$Hue, wine.N$ColorInt, labels = rownames(wine.N), cex = 0.7)
```
Outliers from method 1 - 116, 152, 160, 159, 26, 122, 74, 128, 14, 37, 60

```{r}
#Removed Outliers method 1
o1 <- match(c(116, 152, 160, 159, 26, 122, 74, 128, 14, 37, 60),row.names(wine.N))

cor(wine.N$Alcohol,wine.N$Flavonoids)
cor(wine.N$Alcohol[-o1],wine.N$Flavonoids[-o1])

cor(wine.N$Ash,wine.N$Alk)
cor(wine.N$Ash[-o1],wine.N$Alk[-o1])

cor(wine.N$Hue,wine.N$ColorInt)
cor(wine.N$Hue[-o1],wine.N$ColorInt[-o1])
```
Analysis : In this method, we applied the b

Method 2(Outlier Detector) - Mahalanobis and Chi-Sqr
```{r}
#install.packages("mvoutlier")
library(mvoutlier)
outliers <- aq.plot(wine.N, delta=qchisq(0.975, df = ncol(wine.N)),quan = 1/2)
outliers
```
Outliers from method 2 - 159, 160, 178, 154, 152, 172, 167, 168, 173, 157, 176, 170, 165, 156, 174, 149, 177, 148, 151, 161, 150, 153, 12

```{r}
#Removed Outliers method 2
o2 <- match(c(159, 160, 178, 154, 152, 172, 167, 168, 173, 157, 176, 170, 165, 156, 174, 149, 177, 148, 151, 161, 150, 153, 12),row.names(wine.N))

cor(wine.N$Alcohol,wine.N$Flavonoids)
cor(wine.N$Alcohol[-o2],wine.N$Flavonoids[-o2])

cor(wine.N$Ash,wine.N$Alk)
cor(wine.N$Ash[-o2],wine.N$Alk[-o2])

cor(wine.N$Hue,wine.N$ColorInt)
cor(wine.N$Hue[-o2],wine.N$ColorInt[-o2])
```



Dimensionality & Reduction (PCA & MDS)

New set of data
```{r}
wine.N1 <- wine.N[-c(116, 152, 160, 159, 26, 122, 74, 128, 14, 37, 60),]
wine.N2 <- wine.N[-c(159, 160, 178, 154, 152, 172, 167, 168, 173, 157, 176, 170, 165, 156, 174, 149, 177, 148, 151, 161, 150, 153, 12),]
```

PCA
Outlier 1
```{r}
pca <- princomp(scale(wine.N1))
summary(pca)
plot(pca)
biplot(pca)
```

Outlier 2
```{r}
pca <- princomp(scale(wine.N2))
summary(pca)
plot(pca)
biplot(pca)
```

MDS
Outlier 1 
```{r}
#Scaled distance Matrix
wine.ds <- dist(scale(wine.N1[,1:13]))
round(wine.ds,3)

#Proportion of the Eigenvalues
cmd1 <- cmdscale(wine.ds, k = 7, eig = TRUE)
cumsum(cmd1$eig)/sum(cmd1$eig)

#Multidimensional Scaling
cmd2 <- cmdscale(wine.ds, k = 2, eig = TRUE)
plot(cmd2$points[,1:2], pch = ".")
text(cmd2$points[,1:2], labels = rownames(wine.N1), cex = 0.6)

#Euclidean Distance
wine.ds2 <- dist(wine.N1[,1:7])
cmd3 <- cmdscale(wine.ds2, k = 2, eig = TRUE)
plot(cmd3$points[,1:2], pch = ".")
text(cmd3$points[,1:2], labels = rownames(wine.N1), cex = 0.6)

#Correlation Distance
cor <- cor(wine.N1[,1:13])
cor

cordist <- 1 - cor
cordist

#Multidimensional Scaling of Variables
cmd4 <- cmdscale(cordist, k = 2, eig = TRUE)
plot(cmd4$points[,1:2], pch = ".")
text(cmd4$points[,1:2], labels = rownames(cordist), cex = 0.6)
```

Outlier 2
```{r}
#Scaled distance Matrix
wine.ds <- dist(scale(wine.N2[,1:13]))
round(wine.ds,3)

#Proportion of the Eigenvalues
cmd1 <- cmdscale(wine.ds, k = 7, eig = TRUE)
cumsum(cmd1$eig)/sum(cmd1$eig)

#Multidimensional Scaling
cmd2 <- cmdscale(wine.ds, k = 2, eig = TRUE)
plot(cmd2$points[,1:2], pch = ".")
text(cmd2$points[,1:2], labels = rownames(wine.N2), cex = 0.6)

#Euclidean Distance
wine.ds2 <- dist(wine.N2[,1:7])
cmd3 <- cmdscale(wine.ds2, k = 2, eig = TRUE)
plot(cmd3$points[,1:2], pch = ".")
text(cmd3$points[,1:2], labels = rownames(wine.N2), cex = 0.6)

#Correlation Distance
cor <- cor(wine.N2[,1:13])
cor

cordist <- 1 - cor
cordist

#Multidimensional Scaling of Variables
cmd4 <- cmdscale(cordist, k = 2, eig = TRUE)
plot(cmd4$points[,1:2], pch = ".")
text(cmd4$points[,1:2], labels = rownames(cordist), cex = 0.6)
```

Cluster Analysis

Hierarchical Clustering

Outlier 1
```{r}
#distance matrix for correlation-distance
d <- 1 - cor(wine.N1)
d <- as.dist(d)

#hierarchical clustering and plot of dendrogram
wine.N1.hca <- hclust(d, method = "average")
plot(wine.N1.hca, main = "Hierarchical Clustering Analysis of Gene Data - Average Linkage", xlab = "Sample")

wine.N1.hcs <- hclust(d, method = "single")
plot(wine.N1.hca, main = "Hierarchical Clustering Analysis of Gene Data - single Linkage", xlab = "Sample")

wine.N1.hcc <- hclust(d, method = "complete")
plot(wine.N1.hca, main = "Hierarchical Clustering Analysis of Gene Data - complete Linkage", xlab = "Sample")
```

Outlier 2
```{r}
#distance matrix for correlation-distance
d <- 1 - cor(wine.N2)
d <- as.dist(d)

#hierarchical clustering and plot of dendrogram
wine.N2.hca <- hclust(d, method = "average")
plot(wine.N2.hca, main = "Hierarchical Clustering Analysis of Gene Data - Average Linkage", xlab = "Sample")

wine.N2.hcs <- hclust(d, method = "single")
plot(wine.N2.hca, main = "Hierarchical Clustering Analysis of Gene Data - single Linkage", xlab = "Sample")

wine.N2.hcc <- hclust(d, method = "complete")
plot(wine.N2.hca, main = "Hierarchical Clustering Analysis of Gene Data - complete Linkage", xlab = "Sample")
```

Estimation of Clusters using wgss function

Outlier 1
```{r}
#Scale
wine.N1 <- scale(wine.N1)

#Number of Clusters
plot.wgss <- function(mydata, maxc){
wss <- numeric(maxc)
for (i in 1:maxc){
wss[i] <- kmeans(mydata, centers = i, nstart = 10)$tot.withinss
}
plot(1:maxc, wss, type = "b", xlab = "Number of Clusters",
ylab = "Within Groups Sum of Squares", main = "Scree Plot")
}

plot.wgss(wine.N1,20)
```

Outlier 2
```{r}
#Scale
wine.N2 <- scale(wine.N2)

#Number of Clusters
plot.wgss <- function(mydata, maxc){
wss <- numeric(maxc)
for (i in 1:maxc){
wss[i] <- kmeans(mydata, centers = i, nstart = 10)$tot.withinss
}
plot(1:maxc, wss, type = "b", xlab = "Number of Clusters",
ylab = "Within Groups Sum of Squares", main = "Scree Plot")
}

plot.wgss(wine.N2,20)
```

Finding the K-means for appropriate number of Clusters

Outlier 1
```{r}
km.t <- kmeans(wine.N1, centers = 3)
km.t$tot.withinss
km <- kmeans(wine.N1, centers = 3, nstart = 10)
km$tot.withinss
wine.kmc <- kmeans(wine.N1, centers = 3 , nstart = 10) #Change centers to approciate number
wine.kmc$cluster

pca <- princomp(wine.N1) # How can we characterize the first and second principal components  
plot(pca$scores[,1:2], col = km$cluster)
table(km$cluster)

#Interpreation of Clusters
round(wine.kmc$centers,3) #Change number in respect to clusters
```

Outlier 2
```{r}
km.t <- kmeans(wine.N1, centers = 3)
km.t$tot.withinss
km <- kmeans(wine.N1, centers = 3, nstart = 10)
km$tot.withinss
wine.kmc <- kmeans(wine.N1, centers = 3 , nstart = 10) #Change centers to approciate number
wine.kmc$cluster

pca <- princomp(wine.N1) # How can we characterize the first and second principal components  
plot(pca$scores[,1:2], col = km$cluster)
table(km$cluster)

#Interpreation of Clusters
round(wine.kmc$centers,3) #Change number in respect to clusters
```

Model-Based Clustering of Wine Data

Outlier 1
```{r}
wine.N1 <- wine[-c(116, 152, 160, 159, 26, 122, 74, 128, 14, 37, 60),]
library(mclust)
wine.mbc <- Mclust(wine.N1[,2:14], 3)
table(wine.mbc$classification, wine.N1$Region)
```
```{r}
#Model-Based Clustering Uncertainty
plot(wine.mbc, what = "classification", dimens = c(1,13))
wine.mbc$classification

plot(wine.mbc, what = "uncertainty", dimens = c(1,13))
wine.mbc$uncertainty
```

Outlier 2
```{r}
wine.N2 <- wine[-c(159, 160, 178, 154, 152, 172, 167, 168, 173, 157, 176, 170, 165, 156, 174, 149, 177, 148, 151, 161, 150, 153, 12),]
library(mclust)
wine.mbc <- Mclust(wine.N2[,2:14], 3)
table(wine.mbc$classification, wine.N2$Region)
```
```{r}
#Model-Based Clustering Uncertainty
plot(wine.mbc, what = "classification", dimens = c(1,13))
wine.mbc$classification

plot(wine.mbc, what = "uncertainty", dimens = c(1,13))
wine.mbc$uncertainty
summary(wine.mbc)
```


EFA - CFA

```{r}
wine.N1 <- wine.N[-c(116, 152, 160, 159, 26, 122, 74, 128, 14, 37, 60),]
wine.N2 <- wine.N[-c(159, 160, 178, 154, 152, 172, 167, 168, 173, 157, 176, 170, 165, 156, 174, 149, 177, 148, 151, 161, 150, 153, 12),]
```

Outlier 1
```{r}
#To Perform CFA
#install.packages("sem")
library(sem)

#To Create cfa model
#install.packages("lavaan")
library(lavaan)

#install.packages("semPlot")
library(semPlot)

wine.efa <- factanal(wine.N1,factors = 3)
wine.efa
load <- wine.efa$loadings[,1:2]
plot(load)
text(load,labels=names(wine.N1),cex=.7)
wine.efa$loadings
print(wine.efa$loadings,cut = 0.5)



model <- 'Constituents =~ TotalPhenoles + a*Flavonoids + b*Proanthocyanins + c*Hue +e*OD  
       	Flavour =~ Alcohol + a*ColorInt + b*Proline
       	Inorganic Matter =~ Ash + a*Alk  
      	 
TotalPhenoles~~TotalPhenoles
Flavonoids~~Flavonoids
Proanthocyanins~~Proanthocyanins
OD~~OD
Alcohol~~Alcohol
ColorInt~~ColorInt
Proline~~Proline
Hue~~Hue
Ash~~Ash
Alk~~Alk
NonFL~~NonFL
'
wine.s <-scale(wine.N1)
fit <- sem(model,data = wine.s, orthogonal=TRUE)
summary(fit, fit.measures=TRUE, standardized=TRUE)
fitMeasures(fit,c("agfi",'gfi','srmr','pvalue'))
semPaths(fit,what="std", edge.color="black")
```

Outlier 2
```{r}
#To Perform CFA
#install.packages("sem")
library(sem)

#To Create cfa model
#install.packages("lavaan")
library(lavaan)

#install.packages("semPlot")
library(semPlot)

wine.efa <- factanal(wine.N2,factors = 3)
wine.efa
load <- wine.efa$loadings[,1:2]
plot(load)
text(load,labels=names(wine.N2),cex=.7)
wine.efa$loadings
print(wine.efa$loadings,cut = 0.5)



model <- 'Constituents =~ TotalPhenoles + a*Flavonoids + b*Proanthocyanins + c*Hue +e*OD  
       	Flavour =~ Alcohol + a*ColorInt + b*Proline
       	Inorganic Matter =~ Ash + a*Alk  
      	 
TotalPhenoles~~TotalPhenoles
Flavonoids~~Flavonoids
Proanthocyanins~~Proanthocyanins
OD~~OD
Alcohol~~Alcohol
ColorInt~~ColorInt
Proline~~Proline
Hue~~Hue
Ash~~Ash
Alk~~Alk
NonFL~~NonFL
'
wine.s <-scale(wine.N2)
fit <- sem(model,data = wine.s, orthogonal=TRUE)
summary(fit, fit.measures=TRUE, standardized=TRUE)
fitMeasures(fit,c("agfi",'gfi','srmr','pvalue'))
semPaths(fit,what="std", edge.color="black")
```